HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-03-18 paper/main.tex
edits

These problems are partially addressed by measures that decompose over parts of the evaluated
translation. For automatic measures,
these are often words or n-grams, for manual measures some structural
information is taken into account \parcite{machacek:bojar:segranks:2015}.
% the following is specific to automatic measures:
%quantifying the overlap of its sub-parts relative a reference translation.
A promising line of research decomposes
%More recent work on MT evaluation proposed measures
%that decompose
over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure; the most notable of these measures is
HMEANT \cite{lo2011structured}.
\XXX{Do we want to cut the following shorter?}
However, 
their focus on verb argument structures misses many semantic phenomena
(such as inter-clause linkage and nominal argument structures), providing
only
a limited perspective on the semantic similarity between the translation and reference.
Also, assigning semantic structure to an MT output, especially a poor one,
is difficult and arguably ill-defined.

@@

\begin{compactitem}
\item The components of the translation are ordered differently from the source,
  such that the relation between them changes. For instance, if
  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
  the semantic unit corresponding to the Scene will be marked as Bad, while
  the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
\item A superfluous word or phrase has been inserted into
  the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
\item One of the sub-units is missing from the translation.
  For instance, if ``dog bites man'' is translated into ``dog bites''. \bh{Note that in this case we also 
  get an error in the atomic unit, assuming that ``man'' is a semantic unit in the source, but for the other
  two cases there is no atomic unit error.}
\end{compactitem}

@@

Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
Where $a(u)$ is discontiguous in the text, words between the left
and right boundaries of $a(u)$ which are not contained in it (intervening words)
are presented in a smaller red font.
Intervening words are likely change the meaning of the translation
of $u$, and thus should be paid attention to when considering whether the translation
is correct or not.

@@

In order to compute Kappa, we consider the annotation task as a classification task on 
nodes, and compute agreement on nodes which have two annotations from different annotators.
We report the overall Kappa, as well as separate Kappas on atomic
nodes (annotated as Red, Orange or Green) and the structural nodes (annotated
as Adequate or Bad).
As expected and confirmed by confusion matrices in \figref{fig:heatmap}, there
is generally little confusion between the two types of nodes.

>>>>

These problems are partially addressed by measures that decompose over parts of the evaluated
translation. For automatic measures,
these are often words or n-grams, for manual measures some structural
information is taken into account \parcite{machacek:bojar:segranks:2015}.
% the following is specific to automatic measures:
%quantifying the overlap of its sub-parts relative a reference translation.
A promising line of research decomposes
%More recent work on MT evaluation proposed measures
%that decompose
over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure; the most notable of these measures is
HMEANT \cite{lo2011structured}.
% SAVING SPACE:
% \XXX{Do we want to cut the following shorter?}
% However, 
% their focus on verb argument structures misses many semantic phenomena
% (such as inter-clause linkage and nominal argument structures), providing
% only
% a limited perspective on the semantic similarity between the translation and reference.
% Also, assigning semantic structure to an MT output, especially a poor one,
% is difficult and arguably ill-defined.

@@

\begin{compactitem}
\item The components of the translation are ordered differently from the source,
  such that the relation between them changes. For instance, if
  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
  the semantic unit corresponding to the Scene will be marked as Bad, while
  the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
\item A superfluous word or phrase has been inserted into
  the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
\item One of the sub-units is missing from the translation.
  For instance, if ``dog bites man'' is translated into ``dog bites''.
% XXX let's leave the reviewers to notice this:
%  \bh{Note that in this case we also 
%  get an error in the atomic unit, assuming that ``man'' is a semantic unit in the source, but for the other
%  two cases there is no atomic unit error.}
\end{compactitem}

@@

Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
Where $a(u)$ is discontiguous in the text, words between the left
and right boundaries of $a(u)$ which are not contained in it (intervening words)
are presented in a smaller red font.
Intervening words are likely to change the meaning of the translation
of $u$, and thus should be paid attention to when considering whether the translation
is correct or not.

@@

\tabref{tab:iaa} reports the number of nodes which have two annotations from
different annotators and the corresponding Kappas.
%
%In order to compute Kappa, we consider the annotation task as a classification task on 
%nodes, and compute agreement on nodes which have two annotations from different annotators.
We report the overall Kappa, as well as separate Kappas on atomic
nodes (annotated as Red, Orange or Green) and structural nodes (annotated
as Adequate or Bad).
As expected and confirmed by confusion matrices in \figref{fig:heatmap}, there
is generally little confusion between the two types of nodes.
